---
title: "stage 3 - time series"
author: "Fahad BinThabit"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data set is the total population since 1960 until 2017 
## i am applying several modules , but i will stick with ARIMA to make the prediction 
## all stages are explained 


```{r loading library and dataset}

#load the forecasting package 
library(fpp2)
library(tidyverse)
library(readr)
#load the data 
data <- read_csv("population_total_long.csv")



```

## Declare this as time series data 


```{r pressure, echo=FALSE}
Y <- ts(data[,3],start=c(1960),end=c(2017),frequency = 12)

```

########################
## Prelimiary Analysis
########################



```{r}
#time plot
autoplot(Y) + ggtitle(" world population per country") + ylab("")
```

```{r}
#the data has strong trend . investigate transformation 

# take the first difference in the data to remove the trends 

DY <- diff(Y)
autoplot(DY) + ggtitle(" world population per country") + ylab("")

```

```{r}
# series apears stationary , use to investigate seasonality 

ggseasonplot(DY) + ggtitle("seasonal plot : Change in population") +
 ylab("")

```

##############################
# our series Y has trend and seasonality , to remove the first difference . 
#the differened series still has seasonality
#
# Forecast with various methods
################################

```{r}
# use benchmark method 
#let's use the seasonal naive method as our benchmark


fit <- snaive(DY) #Residual sd: 118614368.4613 
print(summary(fit))
checkresiduals(fit)


```

############################
#try different module
########
# Fit ETS method
#######

```{r}
fit_ets <- ets(Y) #sigma:  sigma:  3.1989
print(summary(fit_ets))
checkresiduals(fit_ets)
```
## fit in ARIMA module 

```{r}
fit_arima <- auto.arima(Y, d=1, D=1, stepwise = FALSE, approximation = FALSE, trace = TRUE)
print(summary(fit_arima))
checkresiduals(fit_arima)
```
#########

#forecast with ARIMA model

```{r}
fcst <- forecast(fit_arima,h=24)
autoplot(fcst, include=90)
print(summary(fcst))
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
